Choosing the Choice: Reflections on Modelling Decisions and Behaviour in Demographic Agent-Based Models

### 2017-03-27 article.tex
attempt to address editor comment 3

This assumed sensitivity of demographic decision-making to the horizons over which the resultant costs and benefits are realised suggests that we should at least acknowledge what assumptions we make about how people value present and future rewards. The flexibility of agent-based models suggests that are is also scope to examine the effect of decision models including time preference. A brief summary of some current thinking on time-dependence in decision making follows; including such models in demographic models will certainly be challenging, and in many cases impossible, but being aware of the potential problems with making simplifying assumptions about treatment of time in decision models is valuable, and at least helps a modeller anticipate when such assumptions may fall down and why.

@@

A crucial point is how much of these empirical and theoretical considerations with respect to time should be included in simulation in a particular case. It may be that the types of simplifications used with respect to time-dependent decisions in many agent-based models are sufficient to capture the essence of the process at hand. However, we would suggest that given that many demographic decisions have long term consequences, an awareness of the choices one is making about time and decision making is, we suggest, invaluable, and testing of different models of time-preference should be considered by modellers if plausible.

@@

All of these problems are far from trivial, particularly as there is little consensus about which models of decision are best in which contexts. How then should we proceed with modelling when the fundamental questions about how individuals make decisions, how they treat time, and how they differ between themselves in these factors are still very uncertain? One way forward is to use the same simulation set-up to analyse multiple models of behaviour \citep{Grimm2005, Epstein2013, Rossiter2014}. This multi-model approach allows the modeller to attempt to distinguish between more and less plausible models of behaviour in the particular research context . Conditional on the simulation set-up being appropriate, more plausible models may be better able to reproduce multiple empirical patterns at varying scales \citep{Werker2004, Bianchi2008}.

>>>>

This assumed sensitivity of demographic decision-making to the horizons over which the resultant costs and benefits are realised suggests that we should at least acknowledge what assumptions we make about how people value present and future rewards. The flexibility of agent-based models suggests that are is also scope to examine the effect of decision models that include time preference. Considering migration models, agents may learn about the potential rewards form migration from their social environment, and such information will influence a prediction about their future conditional on their migration decision. The influence of this prediction on the final decision may then depend on the time horizon over which rewards are realised.  A brief summary of some current thinking on time-dependence in decision making follows; including such models in demographic models will certainly be challenging, and in many cases impossible, but being aware of the potential problems with making simplifying assumptions about treatment of time in decision models is valuable, and at least helps a modeller anticipate when such assumptions may fall down and why.

@@

A crucial point is how much of these empirical and theoretical considerations with respect to time should be included in simulation in a particular case. A model of migration may include exponential or hyperbolical discounting of future income streams associated with migration, as is common in many analytic models \citep{Carrington1996}. This may be crucial to understanding behaviour, as perhaps an adjustment period of obtaining country-specific human captial leads to initially poor earnings \citep{Dustmann2011}, and so the pay-off to migration is disproptionately located at a longer time horizon. Alternatively, as a simplification, agents may be constructed to only consider a choice between current earnings and earnings that could be achieved on arrival. It may be that this type of simplification with respect to time-dependent decisions is sufficient to capture the essence of the process at hand. However, in general we would suggest that given that many demographic decisions have long term consequences, an awareness of the effect of the choices one is making about time and decision making is, we suggest, invaluable, and testing of different models of time-preference should be considered by modellers.

In order to allow for discounting of future reward streams, we must allow for agents to predict their future earnings and/or ability to save conditional on migration. Of course, in practice, predictions of such quantities are rarely perfect and can easily turn out incorrectly \citep{Dustmann1997}. \cite{Sozou1998} describes how the hyperbolic discounting scheme can be derived from an exponential scheme where the reward may be lost in the intervening time period between intial consideration of the choice and realisation of the outcome; as the agent becomes more certain about their eventual receipt of any given reward, the attractiveness of the option increases. The possibility for error in prediction is just one of the many sources of uncertainty surrounding the agent's decision-making process and the attempt to model it.

\subsection{Uncertainty}

@@
